Adaption of the Chumbley Score to matching of bullet striation marks

### 2017-12-26 doc/paper.Rmd
small edits

```{r data-profiles}
if (!file.exists("../data/all-profiles.rds")) {
  files <- dir("../data/profiles/", pattern="csv")
  all <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/profiles", file)) 
    #print(names(tmp))
    if (length(grep("coarse",names(tmp))) == 0)
      tmp$coarse <- gsub(".*coar-(.*).csv", "\\1", file)
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
    all <- rbind(all, tmp)
  }
  all <- all %>% filter(land1_id < land2_id)
  all <- all %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
  saveRDS(all, file="../data/all-profiles.rds")
} else {
  all <- readRDS("../data/all-profiles.rds")
}

errors2 <- rbind(errorrate(all, 0.001), 
                errorrate(all, 0.005), 
                errorrate(all, 0.01), 
                errorrate(all, 0.05))
```

@@

Striation marks are used to determine same source for bullets. Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner. One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citep{NAS:2009}. The issue of subjectivity of a firearms examiner has been reviewed by many authors where things like scientific principle testability and error rates are considered to be defining aspects for an objective analysis such that without identifying the uncertainities associated with a method of comparison, it is inconclusive to regards an analysis complete. This means that with the usual method of visual comparison that is adopted by firearms examiners, lies the problem of lack of any systemized quantification and as such there seems no certain way of associating it with any kind of mathematical probability. Since determining error rates has been duly noted as a fundamental problem in forensic science \citep{NAS:2009}, there is a need of methods that address this problem. Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citep{riva}, although as noted by \citep{aoas}, in this case alignments of striae involves roation of planes, which cannot be generalized for bullets. In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citep{efron}, but there are alternate error estimation techniques as described by \citep{aoas}, \citep{efron} and  \citep{vorburger2016} which give better results.

@@

An important aspect of same source matching in firearms is identification of bullet microtopographies that are unique to be chosen for a match, Bullet profiles and signatures prove to be such features. The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land which minimal noise and many pronounced striation marks. \citet{aoas}

@@

\citep{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons. \citep{hadler} on the other hand, in the improved version of the algorithm first described by \citep{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/B6-B2-L6.png}
```{r, echo=FALSE, fig.width=8, fig.height=3, out.width='\\textwidth'}
land <- read.csv("../data/b6b2l6.csv")
land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw()
```

@@

###Optimization step
The idea behind this step is to first identify the area of best agreement in the two toolmark data. Comparison window size is predefined by the user, which in case of screwdriver toolmarks was chosen by Chumbley et al and Hadler et al as around 10 percent of the length of the toolmarks, which was around 500. This window is henceforth referred to as Window of optimization.

A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what we intuitively expect for matches, and something which we do not intuitively expect for non matches \gk{done}\hh{XXX get rid of the parentheses - put the content into a sentence after your first statement.}

@@

```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair."}
a<- matrix(NA, ncol = 3, nrow = 2)
toolmark.falsepositive<- data.frame(a)
colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
toolmark.falsepositive$Classification<- c("Match", "Non-Match")
toolmark.falsepositive$Match<- c(47,0)
toolmark.falsepositive$`Non-Match`<- c(3,50)
toolmark.falsepositive

@@

## Testing setup:
\hh{Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly}

@@

## Signatures

@@

```{r, echo=FALSE, results='asis'}
param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
#param_setting<- data.frame(matrix(param_setting, nrow = 16))
#t(param_setting)
#param_setting %>% tally() 
library(kableExtra)
knitr::kable(t(param_setting), digits=0, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down"))
```

@@

>>>>

```{r data-profiles}
if (!file.exists("../data/all-profiles.rds")) {
  files <- dir("../data/profiles/", pattern="csv")
  allprofiles <- data.frame()
  for (file in files) {
    tmp <- read.csv(file=file.path(path="../data/profiles", file)) 
    #print(names(tmp))
    if (length(grep("coarse",names(tmp))) == 0)
      tmp$coarse <- gsub(".*coar-(.*).csv", "\\1", file)
    if (length(grep("x",names(tmp)) > 0)) tmp <- tmp %>% select(-x)
    allprofiles <- rbind(allprofiles, tmp)
  }
  allprofiles <- allprofiles %>% filter(land1_id < land2_id)
  allprofiles <- allprofiles %>% mutate(coarse = as.numeric(gsub("pt", ".", coarse)))
  saveRDS(allprofiles, file="../data/all-profiles.rds")
} else {
  allprofiles <- readRDS("../data/all-profiles.rds")
}

errors2 <- rbind(errorrate(allprofiles, 0.001), 
                errorrate(allprofiles, 0.005), 
                errorrate(allprofiles, 0.01), 
                errorrate(allprofiles, 0.05))
```

@@

Striation marks are used to determine same source for bullets. Current standards ask for a visual inspection of the marks under a comparison microscope by a firearms examiner. One of the problems raised by the NAS report in 2009 is the subjectivity involved in this comparison and the need for determining error rates \citep{NAS:2009}. The issue of subjectivity of a firearms examiner has been reviewed by many authors where things like scientific principle testability and error rates are considered to be defining aspects for an objective analysis such that without identifying the uncertainities associated with a method of comparison, it is inconclusive to regards an analysis complete. This means that with the usual method of visual comparison that is adopted by firearms examiners, lies the problem of lack of any systemized quantification and as such there seems no certain way of associating it with any kind of mathematical probability. Since determining error rates has been duly noted as a fundamental problem in forensic science \citep{NAS:2009}, there is a need of methods that address this problem. Some of the methods that have been used to estimate error rates in firearm examinations include Automatic cartridge case comparison where same source identification of cartridge cases and associated error rates have been provided by \citep{riva}, although as noted by \citep{aoas}, in this case alignments of striae involves a rotation of planes, which cannot be generalized for bullets. In case of methods that use machine learning algorithms some methods which are bootsrap based methods often tend to give over-estimated error rates \citep{efron}, but there are alternate error estimation techniques as described by \citep{aoas}, \citep{efron} and  \citep{vorburger2016} which give better results. \hh{XXX give examples for these techniques}

@@

An important aspect of same source matching in firearms is identification of bullet microtopographies that are unique to be chosen for a match, Bullet profiles and signatures prove to be such features. The choice of a bullet signature to uniquely define a bullet depends on finding first, a stable region on a Bullet land with minimal noise and pronounced striation marks \citep{aoas}.

@@

\citep{chumbley} found that the algorithm gives false-positives in slightly over 2 cases (or 2%) of the time and about 9 false-negatives (or 9%) for every 100 comparisons. \citep{hadler} on the other hand, in the improved version of the algorithm first described by \citep{chumbley} found false-positives for 0 cases and 3 false-negatives (or 6%) for 50 comparisons of knwon matches.
\begin{figure}
\centering
\includegraphics[width=\textwidth]{images/B6-B2-L6.png}
```{r, echo=FALSE, fig.width=8, fig.height=3, fig.align="center", out.width='.9\\textwidth'}
land <- read.csv("../data/b6b2l6.csv")
land %>% ggplot(aes(x = y, y=resid)) +geom_line() +theme_bw() +
  theme(axis.text.y = element_blank(), axis.title.y = element_blank())
```

@@

### Optimization step
The idea behind this step is to first identify the area of best agreement in the two toolmark data. Comparison window size is predefined by the user, which in case of screwdriver toolmarks was chosen by \citet{chumbley} and \citet{hadler} as around 10 percent of the length of the toolmarks, which was around 500. \hh{XXX it's not clear that the 500 is referring to the 10 percent.} This window is henceforth referred to as Window of optimization.

A maximum correlation statistic is used to identify the region of best agreement, with the maximum usually seen being near 1 for both cases which is what we intuitively expect for matches, and something which we do not intuitively expect for non matches \gk{done} \hh{XXX get rid of the parentheses - put the content into a sentence after your first statement.}

@@

\hh{XXX Figure~\ref{win-comparison} is an orphan - cite it in the correct place.}
\hh{XXX let's include two or three more windows. We won't be able to use color - so we need to think of  a different way to distinguish between optimization and validation and link corresponding validation windows. Also include the code in the folder on the repo that generates sigs\_generate\_windows.rds }

```{r win-comparison, fig.width=8, fig.height=5, out.width='\\textwidth', fig.cap=" The two plots on the left show how the same shift behaves in case of a matching pair and the two plots on the right show how the different shift behaves in case of a matching pair. "}
a<- matrix(NA, ncol = 3, nrow = 2)
toolmark.falsepositive<- data.frame(a)
colnames(toolmark.falsepositive)<- c("Classification", "Match", "Non-Match")
toolmark.falsepositive$Classification<- c("Match", "Non-Match")
toolmark.falsepositive$Match<- c(47,0)
toolmark.falsepositive$`Non-Match`<- c(3,50)
toolmark.falsepositive

@@

# Testing setup:
\hh{Using cross-validation setup to identify appropriate parameter settings for (a) signatures and (b) profiles directly}

@@

## Study setup

@@

```{r, echo=FALSE, results='asis'}
param_setting<- all %>% select(wo, wv) %>% group_by(wo, wv) %>% tally() %>% select(wo,wv)
#param_setting<- data.frame(matrix(param_setting, nrow = 16))
#t(param_setting)
#param_setting %>% tally() 
library(kableExtra)
knitr::kable(t(param_setting), digits=10, caption = "Overview of parameter settings used for optimization and validation windows for bullet land signatures.", format = "latex" ,booktabs=TRUE) %>% kableExtra::kable_styling(latex_options = c("scale_down"))
```

@@

## Results of Chumbley score
- Nominal alpha value shows dependency on the size of the window of optimization
- Test Fail depends on whether known-match or known non-matches has predictive value
- Type II error least bad for window of validation 30 and window of optimization 120

# Algorithm Modification

The proposed algorithm modification is at the same shift step which comes after the shift distance is identified by finding two windows in the two markings that have the highest correlation.

The modification allows for a "wiggle" room in the second marking of the two sets of markings being compared. This means that each set of same shift windows that are being compared, the second marking will window that is under consideration is allowd to move a little towards the left and a little towards the right.

This gives us a new set of 4 windows, 2 to the left and 2 to the right of the original comparison window. Then the correlations for each one of these 5 windows with the window under consideration of the 1st marking is retrieved (from the validation step correlation matrix) or computed.

Then the window that has the maximum correlation (from the set of 5 windows in the 2nd marking) with the window of the 1st marking is chosen.

This new maximum correlation is then used to compute the U-statistics using the same method as before.

## Expected advantage of modification

We are expecting that this modification would improve the type I and II errors for the good. Less number of false negatives and false positives.
The reason for this expectation is many a times the bullet markings are not made at the same distances for two or more bullets because of the way it comes out of the barrel.

Therefore rigid same shifts might not necessarily compare the right set of windows. In the same shift step for one set of windows, allowing the window in the 2nd marking to wiggle left and right and finding the best match to the window in the 1st marking, lets us adjust for situations where markings are compressed or elongated.

## Dependence on the delta of new windows from the original same shift window

From initial tests the amount of movement to get new windows left and right of the original same shift window seems to directly influence the
the pvalue and U statistic that we look at.

### other questions
How much movement is reasonable and should be used? Is there a way to understand if the 2nd marking in comparison to the 1st marking is compressed or elongated or neither? Should the amount of movement selected depend on this compression or elongation? or using a static delta movement value justifieable.
